### 2018-06-19 references.tex
references.tex - contains links to all important papers

\documentclass[12pt,a4paper,final]{article}
\usepackage[latin1]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\title{References}
\begin{document}
\begin{enumerate}
    \item On AISync
    \begin{itemize}
        \item ZhaMot2017 
        \url{http://iopscience.iop.org/article/10.1088/1361-6544/aa8fe7/meta}

        \item ZhaNisMot2017 
        \url{http://iopscience.iop.org/article/10.1088/1361-6544/aa8fe7/meta}
    \end{itemize}
\end{enumerate}
\end{document}